---
output: pdf_document
---
# pyOpenCL and pyCUDA performance data for gap extension

```{r, echo=FALSE}
## Gets the number of input sequences and sums the total time taken
get_sw_time <- function(fname) {
  alignments <- strsplit(fname, '.', fixed = TRUE)[[1]][2]
  alignments <- as.numeric(strsplit(alignments, '_', fixed = TRUE)[[1]][2])
  data <- readLines(fname)
  times <- data[grep('Time spent on Smith-Waterman', data)]
  times <- as.numeric(sub(".* > ", "", times))
  return(c(alignments, sum(times)))
}
```

```{r, echo=FALSE}

## CPU - command 1 - OpenCL 8 cores
cpu_flist <- list.files('.', pattern='*CPU.log.txt')
cpu <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
  cpu <- rbind(cpu, get_sw_time(fname))
}

## CPU - command 2 - OpenCL 1 core
cpu_flist <- list.files('.', pattern='*CPUSingle.log.txt')
cpuSingle <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
  cpuSingle <- rbind(cpuSingle, get_sw_time(fname))
}

## CPU - command 3 - OpenCLTest 1 core
cpu_flist <- list.files('.', pattern='*CPUoclSingle.log.txt')
cpuOclSingle <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
  cpuOclSingle <- rbind(cpuOclSingle, get_sw_time(fname))
}

## CPU - command 4 - OpenCLTest 8 core
cpu_flist <- list.files('.', pattern='*CPUocl.log.txt')
cpuOcl <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
  cpuOcl <- rbind(cpuOcl, get_sw_time(fname))
}

## GPU - command 1 - OpenCL 
gpu_flist <- list.files('.', pattern='*GPUOpenCL.log.txt')
gpuOcl <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in gpu_flist) {
  gpuOcl <- rbind(gpuOcl, get_sw_time(fname))
}

## GPU - command 2 - CUDA 
gpu_flist <- list.files('.', pattern='*GPUCUDA.log.txt')
gpuCuda <- data.frame()
for (fname in gpu_flist) {
  gpuCuda <- rbind(gpuCuda, get_sw_time(fname))
}

## CPU - command 1 - OpenCL 8 cores
cpu_flist <- list.files('.', pattern='*CPU.GE.log.txt')
cpuGE <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
	cpuGE <- rbind(cpuGE, get_sw_time(fname))
}

## CPU - command 2 - OpenCL 1 core
cpu_flist <- list.files('.', pattern='*CPUSingle.GE.log.txt')
cpuSingleGE <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
	cpuSingleGE <- rbind(cpuSingleGE, get_sw_time(fname))
}

## CPU - command 3 - OpenCLTest 1 core
cpu_flist <- list.files('.', pattern='*CPUoclSingle.GE.log.txt')
cpuOclSingleGE <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
	cpuOclSingleGE <- rbind(cpuOclSingleGE, get_sw_time(fname))
}

## CPU - command 4 - OpenCLTest 8 core
cpu_flist <- list.files('.', pattern='*CPUocl.GE.log.txt')
cpuOclGE <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in cpu_flist) {
	cpuOclGE <- rbind(cpuOclGE, get_sw_time(fname))
}

## GPU - command 1 - OpenCL 
gpu_flist <- list.files('.', pattern='*GPUOpenCL.GE.log.txt')
gpuOclGE <- data.frame(Alignments=numeric(), Time=numeric())
for (fname in gpu_flist) {
	gpuOclGE <- rbind(gpuOclGE, get_sw_time(fname))
}

## GPU - command 2 - CUDA 
gpu_flist <- list.files('.', pattern='*GPUCUDA.GE.log.txt')
gpuCudaGE <- data.frame()
for (fname in gpu_flist) {
	gpuCudaGE <- rbind(gpuCudaGE, get_sw_time(fname))
}
header <- c('V1', 'V2')
names(cpu) <- header
names(cpuSingle) <- header
names(cpuOclSingle) <- header
names(cpuOcl) <- header
names(gpuOcl) <- header
names(gpuCuda) <- header

names(cpuGE) <- header
names(cpuSingleGE) <- header
names(cpuOclSingleGE) <- header
names(cpuOclGE) <- header
names(gpuOclGE) <- header
names(gpuCudaGE) <- header
```

```{r echo=FALSE}
library(knitr)

cellUpdates = read.csv("../cellUpdates.csv", sep="\t", header=FALSE, check.names=FALSE)
cellUpdates = cellUpdates[order(cellUpdates[,1]),]
cellUpdates[,2] = cellUpdates[,2]/1e9
# each points consists of 8 query sequences:
cpu = cpu[order(cpu[,1]),]
cpuSingle = cpuSingle[order(cpuSingle[,1]),]
cpuOclSingle = cpuOclSingle[order(cpuOclSingle[,1]),]
cpuOcl = cpuOcl[order(cpuOcl[,1]),]
gpuOcl = gpuOcl[order(gpuOcl[,1]),]
gpuCuda = gpuCuda[order(gpuCuda[,1]),]
# each points consists of 8 query sequences:
cpu[,1] = cpu[,1]*8
cpuSingle[,1] = cpuSingle[,1]*8
cpuOclSingle[,1] = cpuOclSingle[,1]*8
cpuOcl[,1] = cpuOcl[,1]*8
gpuOcl[,1] = gpuOcl[,1]*8
gpuCuda[,1] = gpuCuda[,1]*8

cpu = cbind(cpu, cellUpdates[,2]/cpu[,2])
cpuSingle = cbind(cpuSingle, cellUpdates[,2]/cpuSingle[,2])
cpuOclSingle = cbind(cpuOclSingle, cellUpdates[,2]/cpuOclSingle[,2])
cpuOcl = cbind(cpuOcl, cellUpdates[,2]/cpuOcl[,2])
gpuOcl = cbind(gpuOcl, cellUpdates[,2]/gpuOcl[,2])
gpuCuda = cbind(gpuCuda, cellUpdates[,2]/gpuCuda[,2])


# remove 0 value
cpu = cpu[cpu$V1>0,]
cpuSingle = cpuSingle[cpuSingle$V1>0,]
cpuOclSingle = cpuOclSingle[cpuOclSingle$V1>0,]
cpuOcl = cpuOcl[cpuOcl$V1>0,]
gpuOcl = gpuOcl[gpuOcl$V1>0,]
gpuCuda = gpuCuda[gpuCuda$V1>0,]

# each points consists of 8 query sequences:
cpuGE = cpuGE[order(cpuGE[,1]),]
cpuSingleGE = cpuSingleGE[order(cpuSingleGE[,1]),]
cpuOclSingleGE = cpuOclSingleGE[order(cpuOclSingleGE[,1]),]
cpuOclGE = cpuOclGE[order(cpuOclGE[,1]),]
gpuOclGE = gpuOclGE[order(gpuOclGE[,1]),]
gpuCudaGE = gpuCudaGE[order(gpuCudaGE[,1]),]
# each points consists of 8 query sequences:
cpuGE[,1] = cpuGE[,1]*8
cpuSingleGE[,1] = cpuSingleGE[,1]*8
cpuOclSingleGE[,1] = cpuOclSingleGE[,1]*8
cpuOclGE[,1] = cpuOclGE[,1]*8
gpuOclGE[,1] = gpuOclGE[,1]*8
gpuCudaGE[,1] = gpuCudaGE[,1]*8

cpuGE = cbind(cpuGE, cellUpdates[,2]/cpuGE[,2])
cpuSingleGE = cbind(cpuSingleGE, cellUpdates[,2]/cpuSingleGE[,2])
cpuOclSingleGE = cbind(cpuOclSingleGE, cellUpdates[,2]/cpuOclSingleGE[,2])
cpuOclGE = cbind(cpuOclGE, cellUpdates[,2]/cpuOclGE[,2])
gpuOclGE = cbind(gpuOclGE, cellUpdates[,2]/gpuOclGE[,2])
gpuCudaGE = cbind(gpuCudaGE, cellUpdates[,2]/gpuCudaGE[,2])


# remove 0 value
cpuGE = cpuGE[cpuGE$V1>0,]
cpuSingleGE = cpuSingleGE[cpuSingleGE$V1>0,]
cpuOclSingleGE = cpuOclSingleGE[cpuOclSingleGE$V1>0,]
cpuOclGE = cpuOclGE[cpuOclGE$V1>0,]
gpuOclGE = gpuOclGE[gpuOclGE$V1>0,]
gpuCudaGE = gpuCudaGE[gpuCudaGE$V1>0,]

```

## Desktop system

These configurations are used:

A) CPU, cpu-OpenCL, 1 core
B) CPU, cpu-OpenCL, 8 cores
C) CPU, gpu-OpenCL, 1 core
D) CPU, gpu-OpenCL, 8 cores
E) GPU, gpu-OpenCL 
F) GPU, CUDA

Definitions:

- GPU: GeForce GTX 650
- CPU: Intel(R) Core(TM) i7-3770 CPU @ 3.40GHz
- CUDA: PaSWAS Smith-Waterman code base
- gpu-OpenCL: for GPU optimized OpenCL code base
- cpu-OpenCL: for CPU optimized OpenCL code base

System: 

- Ubuntu 15.04 
- Desktop PC 


## Performance penalty for gap extension

To show the performance penalty of using the gap extension penalty, each
SW run is compared to the timing of doing a SW with a gap extension penalty. Memory usage, data sets and other setings were kept the same to make sure only the performance penalty of the introduction of the gap extension penalty was measured.

```{r echo=FALSE}
plot(c(0,max(cpuOclSingle$V1)),c(0.5,3), pch='o', cex=0.6, xlab="Number of alignments processed", ylab="Gap extension performance penalty", type="n", xaxt="n")
axis(side=1, at=cpuOclSingle$V1)
points(cpuOclSingle$V1,cpuOclSingleGE$V2/cpuOclSingle$V2, pch='o', cex=0.6)
points(cpuSingle$V1,cpuSingleGE$V2/cpuSingle$V2, col='red', pch='x', cex=0.6)
points(cpuOcl$V1,cpuOclGE$V2/cpuOcl$V2, col='chocolate4', pch='#', cex=0.6)
points(cpu$V1,cpuGE$V2/cpu$V2, col='blue', pch='*', cex=0.6)
points(gpuOcl$V1,gpuOclGE$V2/gpuOcl$V2, col='darkviolet', pch='+', cex=0.6)
points(gpuCuda$V1,gpuCudaGE$V2/gpuCuda$V2, col='darkmagenta', pch='-', cex=0.6)
grid()
legend(0, 3, c("GPU, CUDA",
                  "GPU, gpu-OpenCL",  
                  "CPU, gpu-OpenCL, 8 cores", 
                  "CPU, gpu-OpenCL, 1 core",
                  "CPU, cpu-OpenCL, 8 cores", 
                  "CPU, cpu-OpenCL, 1 core"
), 
col=c('darkmagenta','darkviolet','chocolate4','black','blue','red'), pch=c('-','+','#','o','*','x'), cex=0.6,ncol=3)
```

## Average slow down when using gap extension penalty

| Configuration | Average slow down |
|---------------|-------------------|
| GPU, CUDA | `r mean( gpuCudaGE$V2/gpuCuda$V2)`|
| GPU, gpu-OpenCL | `r mean( gpuOclGE$V2/gpuOcl$V2)`|
| CPU, gpu-OpenCL, 8 cores | `r mean(cpuOclGE$V2/cpuOcl$V2 )` |
| CPU, gpu-OpenCL, 1 core | `r mean( cpuOclSingleGE$V2/cpuOclSingle$V2)` |
| CPU, cpu-OpenCL, 8 cores | `r mean( cpuGE$V2/cpu$V2 )`|
| CPU, cpu-OpenCL, 1 core | `r mean(cpuSingleGE$V2/cpuSingle$V2 )`|
